---
title: "Calendar Snapshot Data gathering V 2"
author: "John D. Smith"
date: "April 28, 2015"
output: html_document
---

Declarations and setup

```{r}
library(RMySQL)
library(stringr)
library(dplyr)
library(lubridate)
library(tidyr)
library(reshape2)
library(xtable)

setwd("~/Documents/Shambhala/PSMC-Comm-IT/statistics/misc-queries")

# con  <- dbConnect(MySQL(), user="Predictor", password= Sys.getenv("SDB_PASSWORD"),
#                   dbname="SDB_STATISTICS", host="46.252.31.8", port=3308)

my_sdb <- src_mysql(user = "Predictor", password = Sys.getenv("SDB_PASSWORD"),
                  dbname = "SDB_STATISTICS", host = "46.252.31.8", port = 3308)
 
source("misc-tables.R")
program_codes_tbl  <- read_tsv("prog_codes_table.tsv")
program_series_tbl <- read_tsv("program_series_table.csv")

```

pull registrations and calendar data

Need to include ALL registrations and the WHOLE shambhala calendar before zeroing in on ONE Center's members / calendar.

```{r, echo=FALSE}

sdb_registrations <- tbl(my_sdb, "registrations")
sdb_registrations_internet <- tbl(my_sdb, "registrations_internet")
sdb_calendar  <- tbl(my_sdb, "calendar")

registrations <- sdb_registrations %>%
  select(reg_id = id, calendar_id = idcalendar, member_id = idmember, registration_date = date) %>%
  collect() %>% 
  mutate(registration_date = sub("0000-00-00 00:00:00", " ", registration_date),  
         registration_date = ymd_hms(registration_date))

registrations_internet <- sdb_registrations_internet %>%
  select(reg_id = id, calendar_id = idcalendar, member_id = idmember, registration_date = date) %>%
  collect() %>% 
  mutate(registration_date = sub("0000-00-00 00:00:00", " ", registration_date),  
         registration_date = ymd_hms(registration_date))

registrations <- bind_rows(registrations, registrations_internet) %>% filter(calendar_id != 0)

registration_stats  <- registrations  %>% group_by(calendar_id)  %>% 
  summarise(first_registrations = min(registration_date, na.rm = T),
            median_registrations = median(registration_date, na.rm = T),
            last_registrations = max(registration_date, na.rm = T),  total_registrations_count = n())

calendar <- sdb_calendar  %>%
  select(calendar_id = id, cal_create_dt = created, date_begin, date_end,  
         program_code = code, program_price = price,
         location, program_name = name, local_event, public, max_reg, 
         teacher1, teacher2, cancelled, visible ) %>% 
  collect() %>% 
    # deal with a bunch of date issues, like missing date, calculate length of time, quarter, etc.
    mutate(
       date_begin = sub("0000-00-00 00:00:00", " ", date_begin),  date_begin = ymd_hms(date_begin),
       date_end = sub("0000-00-00 00:00:00", " ", date_end),  date_end = ymd_hms(date_end),
       cal_create_dt = sub("0000-00-00 00:00:00", " ", cal_create_dt),  cal_create_dt = ymd_hms(cal_create_dt),
       # the following calc messes up date_end, so leaving NA's 
       # date_end = ifelse(is.na(date_end) & !is.na(date_begin), date_begin + 86400, date_end), 
       class_length_days = difftime(date_end, date_begin, units = "days"), 
       day_begin = weekdays(date_begin, abbreviate = F),
       create_to_begin_weeks = difftime(date_begin, cal_create_dt, units = "weeks"),
       quarter_begin = paste( year(date_begin), "-Q", quarter(date_begin), sep = ""),
       program_price_num = sub(" .+$", "", program_price),  
       program_price_num = as.numeric(gsub("\\$", "", program_price_num)),
       weekend_class = ifelse(day_begin %in% c("Friday","Saturday","Sunday") & 
                            class_length_days < 3, 1, 0)
         )
# for historic data, date_end in the calendar table is sometimes missing, 
#   so it would be nice to set it to something reasonable.  Difficult.

calendar <- left_join(calendar, program_codes_tbl)
cal_stats <- left_join(calendar, registration_stats)

# str(calendar)
# str(program_codes_tbl)

```
Pull membership data 
```{r, echo=FALSE}

sdb_members <- tbl(my_sdb, "members")
members  <- sdb_members %>% 
  select(member_id = id, m_rec_date_changed = changed, member_center = center, membertype, 
         m_rec_date_create = created, creator, editor, ismale) %>% 
  collect() %>% mutate(
        m_rec_date_create = sub("0000-00-00 00:00:00", " ", m_rec_date_create),  
        m_rec_date_create = ymd_hms(m_rec_date_create),
        m_rec_date_changed = sub("0000-00-00 00:00:00", " ", m_rec_date_changed),  
        m_rec_date_changed = ymd_hms(m_rec_date_changed))

# Join them all together in various ways

# gather member/non-member counts for each event / center combination:

lj_mem_reg <- left_join(members,registrations) %>% filter(!is.na(reg_id)) %>%
  mutate( ismember = ifelse(membertype %in% c("3", "4","5"), "Member", "Non_member") ) %>% 
  group_by(member_center, ismember, calendar_id) %>% summarise(registrations = n()) %>% 
  select(calendar_id, member_center, ismember, registrations )

# transpose the dataset...
lj_mem_reg_wide <- dcast(lj_mem_reg, calendar_id + member_center ~ ismember, 
                         value.var = "registrations" ) %>% 
  mutate(Member = ifelse(is.na(Member), 0, Member),
         Non_member = ifelse(is.na(Non_member), 0, Non_member),
         total_reg = Member + Non_member)

# gather registration profile for each event.


# est_elders_lang_wide <- dcast(est_elders_lang, country_language+membertype ~ e_older, value.var = "count")
# %>% select(-member_id, -m_rec_date_changed, -m_rec_date_create, -creator, -editor, -reg_id)

dim(lj_mem_reg)
# lj_reg_cal <- left_join(registrations,calendar)
# lj_reg_cal_mem <- left_join(lj_reg_cal, members)

# lj_mem_reg <- left_join(members,registrations)
# lj_mem_reg_cal <- left_join(lj_mem_reg,calendar)
# save(lj_mem_reg_cal,file="lj_mem_reg_cal.Rdata")
```
save dagasets
```{r}
save(cal_stats,calendar,center214,members,registrations,registration_stats,
     file = "all_files.Rdata")

```
working / testing
```{r}

# save(calendar,members,registrations, file = "base_calendaring_files.Rdata")
# save(member_type_tbl,program_codes_tbl,program_series_tbl,type_of_program.tbl,
     # file = "base_calendaring_tables.Rdata")

series_reg <- registrations %>% filter(calendar_id %in% c(55907, 55905, 55906))
series_sample <- left_join(series_reg,calendar)
series_sample %<>% arrange(desc(registration_date)) 
  
series_comp <- dcast(series_sample, calendar_id + program_code + reg_id ~ program_name,value.var =  "registration_date")

nam_list[4]  <- "L01"
nam_list[5]  <- "MEL"
nam_list[6]  <- "SER"
names(series_comp)  <- nam_list

series_comp  %<>% mutate(L01 = as.Date.POSIXct(L01, origin = "1970-01-01 00:00:00 GMT"),
                         MEL = as.Date.POSIXct(MEL, origin = "1970-01-01 00:00:00 GMT"),
                         SER = as.Date.POSIXct(SER, origin = "1970-01-01 00:00:00 GMT")
                         )
series_comp_a <- left_join(series_comp,series_reg) %>% arrange(desc(registration_date))

no_travel_rept_1 <- dcast(no_travel_rept, ever_at_land_center + ever_at_other_center + ever_at_home_center  ~ membertype , value.var = "n_mems")  %>% arrange(desc(ever_at_land_center), desc(ever_at_other_center), desc(ever_at_home_center))


```

Look at individual centers


```{r}

center_id <- "214" # Portland
# center_id <- "211" # Seatttle
# center_id <- "212"  # Austin
# center_id <- "210"  # Philadelphia
# center_id <- "236"  # Vancouver
# center_id <- "196"  # Atlanta
# center_id <- "237"  # Victoria
# center_id <- "211"  # Seattle
# center_id <- "214"  # Portland
# center_id <- "195"  # Minneapolis


# center214 <- lj_reg_cal_mem %>% filter(location == center_id | center == center_id )
center214 <- cal_stats %>% filter(location == center_id & event_type == "Way of Shambhala" )
summary(center214)

# check out https://shambhala.name/report_event.php?id=223092 

#   bg_event2 <- lj_reg_cal_mem %>% filter(calendar_id == 202166)
#   # no location code
#   bg_cal <- calendar %>% filter(calendar_id == 223092)

library(googlesheets)
library(httr)

# 2. Register an application at https://cloud.google.com/console#/project
# Replace key and secret below.
my_google <- oauth_app("google", key = Sys.getenv("jds_shambhala_id"), secret = Sys.getenv("jds_shambhala_secret" ))

google_token <- oauth2.0_token(oauth_endpoints("google"), my_google,
scope = "https://www.googleapis.com/auth/userinfo.profile")

write.csv(center214, "center214.csv", row.names = F) # create local copy
my_ss <- upload_ss("center214.csv") # local copy
file.remove("center214.csv") # local copy

```
Calc based on the calendar itself

  * length of classes (end-date minus start-date)
  * day of week class starts
  * weeks between event first posted and event start date
  * classify as weekend / weeknight
  classify as core curriculum, e.g.,
    event_type in "Introductory Programs", "Sacred Path", "Way of Shambhala"
  classify by program category (core, practice, other)
  *  program price as a number

Depend on member record merge
  % of registrants who are members
  % of registrants who are not local (e.g., out-of-towners)
  weeks between event first posted and first registrant
  median registrant and event start date
  % of your students taking classes elsewhere

teacher info
  % of classes taught by one or more local teachers
  % of classes taught by one or more out-of-towners
  % of classes taught by a local and out-of-towner  ??

After Aggregation
  Average enrollments by program category
  number of public classes
  number of classes with pre-requisites
  % of classes that are not in the core curriculum
  weekend / weeknight ratio

-------
% of classes through Shambhala Online (Rockwell) look like an out of town teacher

```{r}

hasteach  <- calendar  %>% filter(teacher1 > 1)  %>% select(calendar_id, teacher = teacher1)
has2teach  <- calendar  %>% filter(teacher2 > 1)  %>% select(calendar_id, teacher = teacher2)
 # cbind them together, check on whether they are "outsiders" or not, 
 #  then merge back with calendar to classify the event

```

```{r}
cal_stats <- calendar %>% filter(location == center_id, visible == 1, cancelled == 0) %>% 
  group_by(quarter_start) %>% 
  summarize( median = median(create_to_start_weeks, na.rm = T), 
             prog_types = n_distinct(program_code),
             n = n())

cal_stats <- group_by(center214, program_code) %>% summarize(n = n()) %>% 
  arrange(desc(n))

write_csv(cal_stats,"cal_stats.csv")
write_csv(cal_stats,"cal_stats.csv")
  
```
An intermediate level of detail might be:

level	name	series
L01	Level 1	WoS
L02	Level 2	WoS
L03	Level 3	WoS
L04	Level 4	WoS
L05	Level 5	WoS
L06	GES	Sacred Path
L07	Windhorse	Sacred Path
L08	Drala	Sacred Path
L09	Meek	Sacred Path
L10	Perky	Sacred Path
L11 Outrageous/Incscrutable	Sacred Path
L12	Golden Key	Sacred Path
L13	Warrior's Assembly	Sacred Path
TIC	MIEL	IEL Series
LIC	JIEL	IEL Series
GAC	FIEL	IEL Series
DRC	WIEL	IEL Series
BG1	BG1	BG Series
BG2	BG2	BG Series
BG3	BG3	BG Series
RWK	Rigden	Rigden
ESA	ESA	ESA
```