---
title: "Calendar Snapshot V 1"
author: "John D. Smith"
date: "April 28, 2015"
output: html_document
---

Declarations and setup


```{r}
library(RMySQL)
library(stringr)
library(dplyr)
library(lubridate)
library(tidyr)
library(reshape2)
library(xtable)

setwd("~/Documents/Shambhala/PSMC-Comm-IT/statistics/misc-queries")

# con  <- dbConnect(MySQL(), user="Predictor", password= Sys.getenv("SDB_PASSWORD"),
#                   dbname="SDB_STATISTICS", host="46.252.31.8", port=3308)

my_sdb <- src_mysql(user = "Predictor", password = Sys.getenv("SDB_PASSWORD"),
                  dbname = "SDB_STATISTICS", host = "46.252.31.8", port = 3308)
 
source("misc-tables.R")
program_codes_tbl  <- read_tsv("prog_codes_table.tsv")
program_series <- read_tsv("program_series.csv")


```

pull registrations and calendar data

Need to include ALL registrations and the WHOLE shambhala calendar before zeroing in on ONE Center's members / calendar.

```{r, echo=FALSE}

sdb_registrations <- tbl(my_sdb, "registrations")
sdb_registrations_internet <- tbl(my_sdb, "registrations_internet")
sdb_calendar  <- tbl(my_sdb, "calendar")


registrations <- sdb_registrations %>%
  select(reg_id = id, calendar_id = idcalendar, member_id = idmember) %>%
  collect() 

registrations_internet <- sdb_registrations_internet %>%
  select(reg_id = id, calendar_id = idcalendar, member_id = idmember) %>%
  collect() 

registrations <- bind_rows(registrations, registrations_internet) %>% filter (calendar_id != 0)
# for historic data, date_end is sometimes missing, so set it to 1 day after day_start

calendar <- sdb_calendar  %>%
  select(calendar_id = id, cal_create_dt = created, date_begin, date_end,  
         program_code = code, program_price = price,
         location, program_name = name, local_event, public, max_reg, 
         teacher1, teacher2, cancelled, visible ) %>% 
  collect() %>% 
    # deal with a bunch of date issues, like missing date, calculate length of time, quarter, etc.
    mutate(
       cal_create_dt = sub("0000-00-00 00:00:00", " ", cal_create_dt),  
         cal_create_dt = ymd_hms(cal_create_dt),
       date_begin = sub("0000-00-00 00:00:00", " ", date_begin),  date_begin = ymd_hms(date_begin),
       date_end = sub("0000-00-00 00:00:00", " ", date_end),  date_end = ymd_hms(date_end),
       # the following calc messes up date_end, so leaving NA's 
       # date_end = ifelse(is.na(date_end) & !is.na(date_begin), date_begin + 86400, date_end), 
       class_length_days = difftime(date_end, date_begin, units = "days"), 
       day_begin = weekdays(date_begin, abbreviate = F),
       create_to_begin_weeks = difftime(date_begin, cal_create_dt, units = "weeks"),
       quarter_begin = paste( year(date_begin), "-Q", quarter(date_begin), sep = ""),
       program_price_num = sub(" .+$", "", program_price),  
       program_price_num = as.numeric(gsub("\\$", "", program_price_num)),
       weekend_class = ifelse(day_begin %in% c("Friday","Saturday","Sunday") & 
                            class_length_days < 3, 1, 0)
         )

calendar <- left_join(calendar, program_codes_tbl)

str(calendar)
str(program_codes_tbl)

```
Pull membership data 
```{r, echo=FALSE}
# Portland
center_id <- "214"
# Seattle
# center_id <- "211"

# Denver
# center_id <- "218"

# Bellingham
# center_id <- "242"

sdb_members <- tbl(my_sdb, "members")
members  <- sdb_members %>% 
  select(member_id = id, m_rec_date_changed = changed, member_center = center, membertype, 
         m_rec_date_create = created, creator, editor, ismale) %>% 
  collect() %>% mutate(
        m_rec_date_create = sub("0000-00-00 00:00:00", " ", m_rec_date_create),  
        m_rec_date_create = ymd_hms(m_rec_date_create),
        m_rec_date_changed = sub("0000-00-00 00:00:00", " ", m_rec_date_changed),  
        m_rec_date_changed = ymd_hms(m_rec_date_changed))

  # filter(center == center_id) %>% 

# Join them all together in various ways
# lj_mem_reg <- left_join(members,registrations) %>% filter(!is.na(reg_id)) %>% group_by(member_center,membertype,ismale,calendar_id) %>% mutate(registrations = n()) %>% select(-member_id, -m_rec_date_changed, -m_rec_date_create, -creator, -editor, -reg_id)

lj_mem_reg <- left_join(members,registrations) %>% filter(!is.na(reg_id)) %>%
  mutate( ismember = ifelse(membertype %in% c("3", "4","5"), "Member", "Non_member") ) %>% 
  group_by(member_center, ismember, calendar_id) %>% summarise(registrations = n()) %>% 
  select(calendar_id, member_center, ismember, registrations )

lj_mem_reg_wide <- dcast(lj_mem_reg, calendar_id + member_center ~ ismember, value.var = "registrations" ) %>% 
  mutate(Member = ifelse(is.na(Member), 0, Member),
         Non_member = ifelse(is.na(Non_member), 0, Non_member),
         total_reg = Member + Non_member)

         
  
# est_elders_lang_wide <- dcast(est_elders_lang, country_language+membertype ~ e_older, value.var = "count")
# %>% select(-member_id, -m_rec_date_changed, -m_rec_date_create, -creator, -editor, -reg_id)

dim(lj_mem_reg)
# lj_reg_cal <- left_join(registrations,calendar)
# lj_reg_cal_mem <- left_join(lj_reg_cal, members)

# lj_mem_reg <- left_join(members,registrations)
# lj_mem_reg_cal <- left_join(lj_mem_reg,calendar)
# save(lj_mem_reg_cal,file="lj_mem_reg_cal.Rdata")
```

```{r}


```

Look at Portland alone
```{r}


# center214 <- lj_reg_cal_mem %>% filter(location == center_id | center == center_id )
center214 <- lj_reg_cal_mem %>% filter(location == center_id )
summary(center214)

# check out https://shambhala.name/report_event.php?id=223092 

#   bg_event2 <- lj_reg_cal_mem %>% filter(calendar_id == 202166)
#   # no location code
#   bg_cal <- calendar %>% filter(calendar_id == 223092)

```
Calc based on the calendar itself

  * length of classes (end-date minus start-date)
  * day of week class starts
  * weeks between event first posted and event start date
  * classify as weekend / weeknight
  classify as core curriculum, e.g.,
    event_type in "Introductory Programs", "Sacred Path", "Way of Shambhala"
  classify by program category (core, practice, other)
  *  program price as a number

Depend on member record merge
  % of registrants who are members
  % of registrants who are not local (e.g., out-of-towners)
  weeks between event first posted and first registrant
  median registrant and event start date
  % of your students taking classes elsewhere

teacher info
  % of classes taught by one or more local teachers
  % of classes taught by one or more out-of-towners
  % of classes taught by a local and out-of-towner  ??

After Aggregation
  Average enrollments by program category
  number of public classes
  number of classes with pre-requisites
  % of classes that are not in the core curriculum
  weekend / weeknight ratio

-------
% of classes through Shambhala Online (Rockwell) look like an out of town teacher

```{r}

hasteach  <- calendar  %>% filter(teacher1 > 1)  %>% select(calendar_id, teacher = teacher1)
has2teach  <- calendar  %>% filter(teacher2 > 1)  %>% select(calendar_id, teacher = teacher2)
 # cbind them together, check on whether they are "outsiders" or not, 
 #  then merge back with calendar to classify the event

```

```{r}
cal_stats <- calendar %>% filter(location == center_id, visible == 1, cancelled == 0) %>% 
  group_by(quarter_start) %>% 
  summarize( median = median(create_to_start_weeks, na.rm = T), 
             prog_types = n_distinct(program_code),
             n = n())

cal_stats <- group_by(center214, program_code) %>% summarize(n = n()) %>% 
  arrange(desc(n))

write_csv(cal_stats,"cal_stats.csv")
write_csv(cal_stats,"cal_stats.csv")
  
```
An intermediate level of detail might be:

level	name	series
L01	Level 1	WoS
L02	Level 2	WoS
L03	Level 3	WoS
L04	Level 4	WoS
L05	Level 5	WoS
L06	GES	Sacred Path
L07	Windhorse	Sacred Path
L08	Drala	Sacred Path
L09	Meek	Sacred Path
L10	Perky	Sacred Path
L11 Outrageous/Incscrutable	Sacred Path
L12	Golden Key	Sacred Path
L13	Warrior's Assembly	Sacred Path
TIC	MIEL	IEL Series
LIC	JIEL	IEL Series
GAC	FIEL	IEL Series
DRC	WIEL	IEL Series
BG1	BG1	BG Series
BG2	BG2	BG Series
BG3	BG3	BG Series
RWK	Rigden	Rigden
ESA	ESA	ESA
```